---
title: "Jenkins Dashboard"
output: 
  html_document:
    self_contained: no
runtime: shiny
---


A quick look at past investments and current opportunities

## 5-Year History: 2010-2015

This line plot tracks Jenkins investments by focus area over the past 
5 years.  In the past couple of years, the Jenkins Foundations has spent more
on Key Community Health Isues and less on Access to Care.

```{r results='asis', tidy=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
library(plyr)
library(lubridate)
library(reshape2)
library(stringr)
library(googleVis)
library(scales)

op <- options(gvis.plot.tag='chart')

jenkins<-read.csv("jenkins.csv", stringsAsFactors=F)

jenkins$date<-as.POSIXct(jenkins$Grant.Date, format="%m/%d/%Y")
jenkins$year<-year(jenkins$date)

jenkins<-jenkins[!duplicated(jenkins$Grant.Num),]

line5data<-ddply(jenkins, .(j.focus, year), summarize, 
	Total_Investments=sum(Grant.Amt, na.rm=T))
line5data<-dcast(line5data, year~j.focus, value.var="Total_Investments")
line5data$Access.html.tooltip<-paste(line5data$year, "-", dollar(line5data[,2]))
line5data$Issues.html.tooltip<-paste(line5data$year, "-", dollar(line5data[,3]))
line5data$None.html.tooltip<-paste(line5data$year, "-", dollar(line5data[,4]))

line5data<-line5data[,c("year", "Access to Health Care Services", 
                          "Access.html.tooltip", "Key Community Health Issues",
                        "Issues.html.tooltip", "None", "None.html.tooltip")]
line5<-gvisLineChart(line5data, xvar="year", 
                     yvar=c("Access to Health Care Services", 
                          "Access.html.tooltip", "Key Community Health Issues",
                        "Issues.html.tooltip", "None", "None.html.tooltip"),
     options=list( 
		vAxis="{format:'currency'}",
		hAxis="{format:'none'}",
		bar="{groupWidth:'90%'}",
		colors="['#37939B', '#FFC20E', '#B30838']",
    lineWidth=2,
		title={"Investments by Focus Area"}))
plot(line5)
```

## Cumulative grant investments  by  strategy

This bar chart illustrates the Jenkins' foundations cumulative investments in different strategy areas since 2010.  

```{r results='asis', tidy=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=10}
bar5data<-ddply(jenkins, .(j.strategy), summarize, 
	Total_Investments=sum(Grant.Amt, na.rm=T))
bar5data$j.strategy<-as.factor(bar5data$j.strategy)
	bar5data$j.strategy<-reorder(bar5data$j.strategy, 
		bar5data$Total_Investments)
bar5data$Investment.html.tooltip<-dollar(bar5data$Total_Investments)

bar5<-gvisBarChart(bar5data, xvar="j.strategy", yvar=c("Total_Investments", 
                                                        "Investment.html.tooltip"),
     options=list( 
		hAxis="{format:'currency'}",
		bar="{groupWidth:'90%'}",
		colors="['#37939B']",
		title={"Cumulative Investments by Strategy"}))
plot(bar5)
```
## Current Opportunities

### Greater Richmond Community Health Issues
```{r results='asis', tidy=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
library(rgeos)
library(zoo)
library(gridExtra)
library(maps)
library(ggmap)
library(mapproj)
library(rgdal)
library(plotly)
library(maptools)
library(tigris)

apps<-read.csv("jenkins apps.csv", stringsAsFactors=F)

japps1<-apps[apps$Organization.Name %in% jenkins$Ctc.Organization,]
japps2<-apps[str_detect(apps$Results.Statement, "Health"),]

japps<-rbind(japps1, 
  japps2[!(japps2$Organization.Name %in% japps1$Organization.Name),])

# Now handcoding jenkins strategies in excel
#write.csv(japps, "jenkins only apps dump.csv")
# Reading in the new handcoded data

japps<-read.csv("jenkins only apps.csv", stringsAsFactors=F)

#### CREATING DF FOR MAPPING GRANT DATA ####
# first need to read in map data 
censusdata<-readOGR(dsn=".", layer="tl_2015_us_county", verbose=F)
censusdata<-censusdata[censusdata$STATEFP==51,]

censusdata$id<-tolower(censusdata$NAME)
censusdata$proj4string <- "+proj=longlat +ellps=clrk66"

atlas<-read.csv("atlas.csv", stringsAsFactors=F, na.str=c("--", "N/A"))

### Need to clean up foundant data/jenkins applications
### so that they are in the same shape as the GIS/map data

map.data<-japps[ , c("Chesterfield.County", 
	"Colonial.Heights", "Goochland.County", "Hanover.County", 
	"Henrico.County", "Hopewell", "Petersburg", 
	"Powhatan.County", "Richmond.City", "j.strategy",	
	"Amount.Requested")]

foundantDF<-melt(map.data, id.vars=c("j.strategy", "Amount.Requested"), 
	variable.name="county")
foundantDF<-ddply(foundantDF, .(j.strategy, county), summarize, 
	num.served=sum(value, na.rm=T))
castFoundantDF<-dcast(foundantDF, county~j.strategy, value.var="num.served")

castFoundantDF$id<-tolower(castFoundantDF$county)
castFoundantDF$id<-gsub("[.]", " ", castFoundantDF$id)
castFoundantDF$id<-gsub("county", "", castFoundantDF$id)
castFoundantDF$id<-gsub("city", "", castFoundantDF$id)
castFoundantDF$id<-str_trim(castFoundantDF$id, "right")


### Need to clean up the health atlas data
### so that they are in the same shape as the GIS/map data

atlasDF<-melt(atlas, id.vars="indicator", variable.name="county")
atlasDF<-atlasDF[!duplicated(atlasDF),] # some indicators are duplicated
castAtlasDF<-dcast(atlasDF, county~indicator, value.var="value", 
	fun.aggregate=sum)

castAtlasDF$id<-castAtlasDF$county
castAtlasDF$id<-gsub("[.]", " ", castAtlasDF$county)

mapdata<-join(castFoundantDF, castAtlasDF, by="id", type="right")

### Now need to get center points for overlaying points on heat map

chesterfieldLon<-geocode("chesterfield county va")[[1]]
colonialHeightsLon<-geocode("colonial heights, va")[[1]]
goochlandLon<-geocode("goochland county va")[[1]]
hanoverLon<-geocode("hanover county va")[[1]]
hopewellLon<-geocode("hopewell county va")[[1]]
petersburgLon<-geocode("petersburg, va")[[1]]
powhatanLon<-geocode("powhatan county va")[[1]]
richmondLon<-geocode("richmond, va")[[1]]
henricoLon<-geocode("henrico county, va")[[1]]
newKentLon<-geocode("new kent county, va")[[1]]
charlesCityLon<-geocode("charles city county, va")[[1]]



chesterfieldLat<-geocode("chesterfield county va")[[2]]
colonialHeightsLat<-geocode("colonial heights, va")[[2]]
goochlandLat<-geocode("goochland county va")[[2]]
hanoverLat<-geocode("hanover county va")[[2]]
hopewellLat<-geocode("hopewell county va")[[2]]
petersburgLat<-geocode("petersburg, va")[[2]]
powhatanLat<-geocode("powhatan county va")[[2]]
richmondLat<-geocode("richmond, va")[[2]]
henricoLat<-geocode("henrico county, va")[[2]]
newKentLat<-geocode("new kent county, va")[[2]]
charlesCityLat<-geocode("charles city county, va")[[2]]

mapdata$centerLon<-NA
mapdata$centerLon[mapdata$id=="chesterfield"]<-chesterfieldLon
mapdata$centerLon[mapdata$id=="colonial heights"]<-colonialHeightsLon
mapdata$centerLon[mapdata$id=="goochland"]<-goochlandLon
mapdata$centerLon[mapdata$id=="hanover"]<-hanoverLon
mapdata$centerLon[mapdata$id=="hopewell"]<-hopewellLon
mapdata$centerLon[mapdata$id=="petersburg"]<-petersburgLon
mapdata$centerLon[mapdata$id=="powhatan"]<-powhatanLon
mapdata$centerLon[mapdata$id=="richmond"]<-richmondLon
mapdata$centerLon[mapdata$id=="henrico"]<-henricoLon
mapdata$centerLon[mapdata$id=="new kent"]<-newKentLon
mapdata$centerLon[mapdata$id=="charles city"]<-charlesCityLon

mapdata$centerLat<-NA
mapdata$centerLat[mapdata$id=="chesterfield"]<-chesterfieldLat
mapdata$centerLat[mapdata$id=="colonial heights"]<-colonialHeightsLat
mapdata$centerLat[mapdata$id=="goochland"]<-goochlandLat
mapdata$centerLat[mapdata$id=="hanover"]<-hanoverLat
mapdata$centerLat[mapdata$id=="hopewell"]<-hopewellLat
mapdata$centerLat[mapdata$id=="petersburg"]<-petersburgLat
mapdata$centerLat[mapdata$id=="powhatan"]<-powhatanLat
mapdata$centerLat[mapdata$id=="richmond"]<-richmondLat
mapdata$centerLat[mapdata$id=="henrico"]<-henricoLon
mapdata$centerLat[mapdata$id=="new kent"]<-newKentLat
mapdata$centerLat[mapdata$id=="charles city"]<-charlesCityLat


colnames(mapdata)<-gsub(",", ".", colnames(mapdata))
colnames(mapdata)<-gsub("#", "num", colnames(mapdata))
colnames(mapdata)<-gsub("%", "percent", colnames(mapdata))
colnames(mapdata)<-gsub("[(]", ".", colnames(mapdata))
colnames(mapdata)<-gsub("[)]", ".", colnames(mapdata))
colnames(mapdata)<-gsub(" ", ".", colnames(mapdata))

data<-geo_join(censusdata, mapdata, "id", "id") 


#get rid of richmond county
data[which(data$COUNTYFP==159),]<-NA

#get rid of powhatan county
data[which(data$COUNTYFP==145),]<-NA

gauge<-atlas[,2:12]
gauge$regionTotal<-apply(gauge, 1, function(x) sum(x, na.rm=T))
gauge$indicator=atlas$indicator
gauge$regionP<-gauge$regionTotal/973206
gauge$regionP[str_detect(gauge$indicator, "18+")]<-gauge$regionTotal[str_detect(gauge$indicator, "18+")]/852855
gauge$regionP<-round(gauge$regionP*100, 1)
gauge$VAp<-(atlas$virginia/6356993)*100
gauge$regionDif<-gauge$regionP-gauge$VAp


keep<-c( "Fair or Poor Health Status (age 18+, #)",
	"Diabetes (age 18+, #)",
	"High Blood Pressure (age 18+, #)",
	"Oral Health: Access (age 18+ no dental visit in past year, #)",
	"Overweight or Obese (age 18+, #)",
	"Access to doctor (age 18+ could not see doctor due to cost, #)",
	"Uninsured (age 0-64, #)" ,
	"Binge Alcohol Use (age 18+, #)",
	"Teenage Pregnancy (age 10-19, #)" 
	)

# get smaller set of these indicators
gauge<-subset(gauge, gauge$indicator %in% keep & is.na(gauge$regionP)==F)

gauge<-gauge[!duplicated(gauge$indicator),c("regionP", "indicator")]

gauge$labels<-c("No Access to Dr.","Binge Alcohol Use","Diabetes",
                "Poor Health Status","High Blood Pressure", 
	"No Oral Health Acces", "Overweight/Obese", "Teenage Pregnancy")

gauge<-gauge[,c("labels", "regionP")] #string labels have to be first column

gaugePlot<-gvisGauge(gauge, labelvar="labels", numvar="regionP",
			options=list(min=0, max=70, greenFrom=0,
                 greenTo=10, yellowFrom=10, yellowTo=20,
                 redFrom=20, redTo=70, width=1000, height=1000))

plot(gaugePlot)
```

### Strength of Evidence and Evaluation Practices
```{r results='asis', tidy=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
treeData<-data.frame(Organization.Name=c("Strategy", levels(as.factor(japps$j.strategy)),
			japps$Organization.Name),
	j.strategy=c(NA, rep("Strategy", length(levels(as.factor(japps$j.strategy)))), 
			japps$j.strategy),
	Amount.Requested=c(sum(japps$Amount.Requested), rep(sum(japps$Amount.Requested),
		length(levels(as.factor(japps$j.strategy)))), japps$Amount.Requested),
	Evaluation.Score=c(4, rep(4,length(levels(as.factor(japps$j.strategy)))), 
		japps$Is.the.logic.model.complete.and.organized.correctly.)
	)

tree<-gvisTreeMap(treeData, idvar="Organization.Name",
	parentvar="j.strategy", size="Amount.Requested", colorvar="Evaluation.Score")
plot(tree)
```

### Investments by Geography
These maps show how the Jenkins Investments are aligned with community needs.

```{r results='asis', tidy=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
library(leaflet)
library(shiny)

strategies<-c("Primary Care", "Healthy Aging", "Mental Health Care",
            "Substance Abuse Treatment and Prevention", "Oral Health Care",
           "Violence Prevention")
pal <- colorBin("OrRd", NULL, n = 5, pretty=F)
county_popup <- paste0("<strong>County: </strong>", 
                      toupper(data$id))
primaryCareMap<-leaflet(data) %>% 
  setView(lng =  -77.43605, lat = 37.54072, zoom = 10)%>% 
  addTiles() %>%
	addPolygons(
  fillColor = ~pal(Access.to.doctor..age.18..could.not.see.doctor.due.to.cost..num.), 
		fillOpacity = 0.8, 
            color = "#BDBDC3", 
            weight = 1, 
            popup = county_popup) %>%
	addCircles(data=mapdata, lng =~centerLon, lat =~centerLat,
		 weight = 1, radius=~Primary.Care*30, color="#00b3b3",
		 fillOpacity=.5)

healthyAgingMap<-leaflet(data) %>% 
  setView(lng =  -77.43605, lat = 37.54072, zoom = 10)%>% 
	addTiles() %>%
	addPolygons(fillColor = ~pal(Seniors.Age.65...num), 
		fillOpacity = 0.8, 
            color = "#BDBDC3", 
            weight = 1, 
            popup = county_popup) %>%
	addCircles(data=mapdata, lng =~centerLon, lat =~centerLat,
		 weight = 1, radius=~Healthy.Aging*30, color="#00b3b3",
		 fillOpacity=.5)

mentalHealthMap<-leaflet(data) %>% 
  setView(lng =  -77.43605, lat = 37.54072, zoom = 10) %>% 
  addTiles() %>%
	addPolygons(fillColor = ~pal(Behavioral.Health.Hospitalizations.Total..rate), 
		fillOpacity = 0.8, 
            color = "#BDBDC3", 
            weight = 1, 
            popup = county_popup) %>%
	addCircles(data=mapdata, lng =~centerLon, lat =~centerLat,
		 weight = 1, radius=~Mental.Health.Care*30, color="#00b3b3",
		 fillOpacity=.5)

oralHealthMap<-leaflet(data) %>% 
  setView(lng =  -77.43605, lat = 37.54072, zoom = 10)%>% 
  addTiles() %>%
	addPolygons(fillColor = ~pal(Oral.Health..Access..age.18..no.dental.visit.in.past.year..num.), 
		fillOpacity = 0.8, 
            color = "#BDBDC3", 
            weight = 1, 
            popup = county_popup) %>%
	addCircles(data=mapdata, lng =~centerLon, lat =~centerLat,
		 weight = 1, radius=~Oral.Health.Care*30, color="#00b3b3",
		 fillOpacity=.5)

substanceAbuseMap<-leaflet(data) %>% 
  setView(lng =  -77.43605, lat = 37.54072, zoom = 10)%>% 
  addTiles() %>%
	addPolygons(fillColor = ~pal(Binge.Alcohol.Use..age.18...num.), 
		fillOpacity = 0.8, 
            color = "#BDBDC3", 
            weight = 1, 
            popup = county_popup) %>%
	addCircles(data=mapdata, lng =~centerLon, lat =~centerLat,
		 weight = 1, radius=~Substance.Abuse.Treatment.and.Prevention*30, color="#00b3b3",
		 fillOpacity=.5)

violencePreventionMap<-leaflet(data) %>% 
  setView(lng =  -77.43605, lat = 37.54072, zoom = 10) %>% 
  addTiles() %>%
	addPolygons(fillColor = ~pal(Bullying..School.Property..age.14.19.in.past.year..num.), 
		fillOpacity = 0.8, 
            color = "#BDBDC3", 
            weight = 1, 
            popup = county_popup) %>%
	addCircles(data=mapdata, lng =~centerLon, lat =~centerLat,
		 weight = 1, radius=~Violence.Prevention*30, color="#00b3b3",
		 fillOpacity=.5)

selectInput("strategies", "Strategies", strategies)
renderLeaflet({
  if(input$strategies=="Primary Care"){
    primaryCareMap
  }else if(input$strategies=="Healthy Aging"){
    healthyAgingMap
  }else if(input$strategies=="Mental Health Care"){
    mentalHealthMap
  }else if(input$strategies=="Oral Health Care"){
    oralHealthMap
  }else if(input$strategies=="Substance Abuse Treatment and Prevention"){
    substanceAbuseMap
  }else violencePreventionMap
    
})

```



